---
title: "Data Analysis"
author: "Sridhar"
date: "28 January 2019"
output: html_document
---

```{r}
library(ggplot2)
library(ggvis)
```
## Load the Data
Breast Cancer Wisconsin(Diagnosis) data is loaded and the column names are defined based on the attribute information.

### Dataset 1
The outcome variable is the "classes" and it has following category of data
* Malignant or
* Benign breast mass
The phenotypes for characterization are
* Sample ID (code number)
* Clump thickness
* Uniformity of cell size
* Uniformity of cell shape
* Marginal adhesion
* Single epithelial cell size
* Number of bare nuclei
* Bland chromatin
* Number of normal nuclei
* Mitosis
* Classes, i.e. diagnosis


```{r setup, include=FALSE}
diagnosis_data <- read.csv("breast-cancer-wisconsin.data.csv", header = FALSE)
colnames(diagnosis_data) <- c("sample_code_number", "clump_thickness", "uniformity_of_cell_size", "uniformity_of_cell_shape", "marginal_adhesion", "single_epithelial_cell_size", 
                       "bare_nuclei", "bland_chromatin", "normal_nucleoli", "mitosis", "classes")

```

The data is presented below.

```{r pressure, echo=FALSE}
head(diagnosis_data)
```

The properties of data are given,

```{r}
str(diagnosis_data)
```

### Dataset 2
The outcome variable is the "classes" and it has following category of data
Attribute Information:

1) ID number 
2) Diagnosis (M = malignant, B = benign) 
3-32) 

Ten real-valued features are computed for each cell nucleus: 

a) radius (mean of distances from center to points on the perimeter) 
b) texture (standard deviation of gray-scale values) 
c) perimeter 
d) area 
e) smoothness (local variation in radius lengths) 
f) compactness (perimeter^2 / area - 1.0) 
g) concavity (severity of concave portions of the contour) 
h) concave points (number of concave portions of the contour) 
i) symmetry 
j) fractal dimension ("coastline approximation" - 1)

```{r}
diagnosis_data_2 <- read.csv("wdbc.data.csv", header = FALSE)

phenotypes <- rep(c("radius", "texture", "perimeter", "area", "smoothness", "compactness", "concavity", "concave_points", "symmetry", "fractal_dimension"), 3)
types <- rep(c("mean", "se", "largest_worst"), each = 10)

colnames(diagnosis_data_2) <- c("ID", "diagnosis", paste(phenotypes, types, sep = "_"))

# how many NAs are in the data
length(which(is.na(diagnosis_data_2)))
```


## Visualizations

```{r}
p <- ggvis(diagnosis_data, ~normal_nucleoli, y = ~mitosis)
layer_points(p)
```

