---
title: "Breast Cancer Dataset Analysis"
author: "Ravindra Peddinti"
date: "December 05, 2017"
output:
  html_document: 
    theme: lumen
    toc: true
  html_notebook: default
  
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, warning=FALSE)
```

## Summary

This is an analysis of the *Breast Cancer Wisconsin (Diagnostic) DataSet*, obtained from [Kaggle](https://www.kaggle.com/uciml/breast-cancer-wisconsin-data)
Predicting the cancer type (benign,malignant) for the given data attributes  based on neural network



## Data loading and cleaning

```{r, message=F, warning=F}
library(knitr)
library(corrplot)
library(funModeling)
library(gmodels)
library(C50)
library(NeuralNetTools)
library(caret)
library(nnet)
library(partykit)
library(gridExtra)
library(devtools)
```


### Data Load
```{r}
 cancerdata <- read.csv("~/Desktop/BreastCancer/Research/data.csv")
```
                     
### Compactly Display the Structure of  cancerdata

```{r}
str(cancerdata)
```

### Identify columns with NA

```{r}
colnames(cancerdata)[ apply(cancerdata, 2, anyNA) ]
#Column X contains NA values. This column will be limited in the Analysis/Modelling dataset.
```
### limiting  X column for data modelling
```{r}
cancerdata <- subset(cancerdata, select = -c(X))
```

### diagnosis column conversion
```{r}
cancerdata$diagnosis<-as.factor(cancerdata$diagnosis) 
```

## Data Analysis
### Express Table Entries as Fraction of Marginal Table
```{r}
  prop.table(table(cancerdata$diagnosis))
```
### Get a summary for the given data frame 
                
  Overview: Quantity of zeros, NA, unique values; as well as the data type may  lead to a good or bad model. 
  Checking NA, zeros, data type and unique values
  q_zeros: quantity of zeros (p_zeros: in percentage)
  q_na: quantity of NA (p_na: in percentage)
  type: factor or numeric
  unique: quantity of unique values

```{r, results=F}
   cancerdata_status=df_status(cancerdata)
```
### Plotting numerical data
```{r}
  plot_num(cancerdata)
```

##visualization of a correlation matrix.

```{r}
corrplot <- cor(cancerdata[,3:ncol(cancerdata)])
corrplot(corrplot, order = "hclust", tl.cex = 0.65, addrect = 8)
```


## Modeling  

### Data data partition to build data model 
                   
```{r}
set.seed(1234)

dataindex <- createDataPartition(cancerdata$diagnosis, p=0.8, list = FALSE)   # 80:20 data partition
 train_data <- cancerdata[dataindex, -1]         # 80% train data
 test_data <-  cancerdata[-dataindex, -1]       # Rest 20% validation data
```

## decision tree model 
### Create model from train data using decision tree model 
### We now create a basic C5.0 model
```{r}
model<-C5.0(diagnosis ~ ., data=train_data)     # Decision tree model  We use DEFAULT parameter values
```
### List important variables in the model
 ```{r}
 C5imp(model)
 ```
### Make prediction and check accuracy
```{r}
 result<-predict(model, test_data[,-1])           # 1st column is held back. It is diagnosis 
```
```{r}
(accuracy<-sum(result == test_data$diagnosis)/nrow(test_data))        # decision tree Accuracy
```
###   Prediction on validation set
```{r}
 model_nnet<-nnet(diagnosis ~. ,data= train_data, size=10)
 result1<-predict(model_nnet, test_data[,-1], type = c("class") )       # Prediction on validation set
(accuracy<-sum(result1 == test_data$diagnosis)/nrow(test_data))        # nnet Accuracy
```

### CrossTable
```{r}
CrossTable(test_data$diagnosis, result, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c("Actual","Predicted "))
```

### Plot Decision Tree
```{r}
columncol<-hcl(c(270, 260, 250), 200, 30, 0.6)
labelcol<-hcl(200, 200, 50, 0.2)
indexcol<-hcl(150, 200, 50, 0.4)
model_tree=C50:::as.party.C5.0(model)
plot(model_tree,type="simple",gp = gpar(fontsize = 8), drop_terminal = TRUE, tnex=1,
inner_panel = node_inner(model_tree, abbreviate = FALSE,
fill = c(labelcol, indexcol), pval = TRUE, id = TRUE),
terminal_panel=node_barplot(model_tree, col = "black", fill = columncol[c(1,2,4)], 
beside = TRUE, ymax = 1, ylines = TRUE, widths = 1, gap = 0.1,
reverse = FALSE, id = TRUE))
```
                                
                                

    
## Neural Networks
```{r}
 model_nnet <- nnet(diagnosis ~ .,data=train_data,size= 20, decay= 0.01, rang=0.6,trace=TRUE,maxit=200 )
```
### Make predictions for validation dataset

```{r}
##   type="class" gives class_values (1 or 0 )
save(model_nnet , file = 'Neuralnetwork.rda')
class_pred <- predict(model_nnet,test_data[,-1])
print(class_pred)
```


###  Plot neural network
```{r}
rel_imp <- garson(model_nnet, bar_plot = FALSE)$rel_imp
cols <- colorRampPalette(c('lightgreen', 'darkgreen'))(3)[rank(rel_imp)]
plotnet(model_nnet, cex_val = .5,circle_col = list(cols, 'lightblue'))
```

##References

| <a href="https://www.kdnuggets.com/2016/08/begineers-guide-neural-networks-r.html">Neural Networks Begineers guide</a>
| <a href="https://rmanic.wordpress.com/2016/05/19/creating-a-well-formatted-decision-tree-with-partykit-and-listing-the-rules-of-the-nodes/">Decision tree with partykit library</a>
| <a href="https://beckmw.wordpress.com/tag/neural-network/">Visualizing neural networks</a>
| <a href="https://eight2late.wordpress.com/2016/02/16/a-gentle-introduction-to-decision-trees-using-r/">Introduction to decision trees</a>




